Conditional selectivity for Hexastores

### 2011-07-11 conditional/trunk/selectivity-hexastore.tex
commutativity
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@4 43ef1190-90fd-4dd7-a769-d0121f02b80e
unit for unbound
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@5 43ef1190-90fd-4dd7-a769-d0121f02b80e
Hexastore start
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@6 43ef1190-90fd-4dd7-a769-d0121f02b80e

\documentclass[12pt]{article}
\usepackage{cite}

@@

Analogously, the conditional selectivity for a triple $T$ can be written as 
\begin{equation}
sel(T) = sel(S \cap P \cap O) = sel(O | S \cap P) \cdot sel(P | S)
\cdot sel(S) . 
\end{equation}

>>>>

\documentclass[12pt, draft]{article}
\usepackage{cite}

@@

Analogously, the conditional selectivity for a triple $T$ can be written as 
\begin{equation}
sel(T) = sel(S \cap P \cap O) = sel(O | S \cap P) \cdot sel(P | S)
\cdot sel(S) . 
\end{equation}
We note that intersection is commutative, thus similar
relations can be shown for each permutation of nodes.

Finally, note that for unbound nodes (variables), the selectivity will
always be unity, whether conditional or independent, since it will
match all possible RDF nodes.

% TODO example?

\section{Exact selectivity for Hexastore}

The Hexastore was first described in
\cite{Weiss:2008:HSI:1453856.1453965}, which sought to exploit a key
strength of the RDF triple model by using six-way indices for
efficient lookup.

A Hexastore contains the six indices \textsf{spo}, \textsf{pso},
\textsf{osp}, \textsf{sop}, \textsf{pos} and \textsf{ops}. Each index
has the following structure, taking \textsf{spo} as an example: ``A
subject key $s_i$ is associated to a sorted vector of $n_i$ property
keys, $\{p_1^i , p_2^i , \ldots , p^i_{n_i} \}$. Each property key $p_j^i$
is, in its turn, linked to an associated sorted list of $k_{i,j}$
object keys.''\cite{Weiss:2008:HSI:1453856.1453965}, section~4.1.

Now, given a query, we shall explain how the selectivity can be
computed exactly rather than being estimated. First, we define a
function $cm(x)$ which returns the number of matches of a given SPARQL
node $x$. % TODO: needs to be a lot more precise
